Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2020-08-26 paper.Rnw
small changes to references

We apply LIME using six sets of tuning parameters to all observations in the \data \ test set to observe variability across tuning parameters. A quantile bin based simulation method is used for five of the LIME applications with the number of bins varying from 2 to 6 by application. We use 6 bins as the maximum, because the complexity of the explanations increases with the number of bins. The sixth application of LIME uses a kernel density simulation method. The default methods for feature selection (forward selection) and the computation of the weights (Gower distance raised to an exponent of 1) are used for all applications. (See Appendix~\ref{lime-details} for more details on the tuning parameters.)

@@

The largest limitation to the diagnostic visualizations is the dimensionality of the data shown, both in the number of dimensions or features as well as the number of observations. Fortunately, in the situation of LIME, both of these aspects are rather well controlled: LIME relies heavily on simulations to generate data scenarios that are close to the data observed but exhibits variability. Effects from overplotting should be relatively mild, because output from simulations is shown, which is expected to be (relatively) continuous \hh{such that overplotting only occurs for points with (relatively) similar values.} \kgc{What do you mean by continuous here?}. \hh{XXX Does that extra explanation help?} \kgc{Yes. I see what you mean now.} In that respect, the diagnostics shown for the \data\ and in the bullet example are representative of what is expected. But in cases where overplotting does become problematic, the user could either simply reduce the size of the simulations or use some well-studied binning techniques in the visualizations, as discussed for example in \citet{carr:1987} or \citet{gold}.

>>>>

We apply LIME using six sets of tuning parameters to all observations in the \data \ test set to observe variability across tuning parameters. A quantile bin based simulation method is used for five of the LIME applications with the number of bins varying from 2 to 6 by application. We use 6 bins as the maximum, because the complexity of the explanations increases with the number of bins. The sixth application of LIME uses a kernel density simulation method. The default methods for feature selection (forward selection) and the computation of the weights (Gower distance raised to an exponent of 1) are used for all applications. (See Appendix~\ref{lime-details} for descriptions of the tuning parameters.)

@@

The largest limitation to the diagnostic visualizations is the dimensionality of the data shown, both in the number of dimensions or features as well as the number of observations. Fortunately, in the situation of LIME, both of these aspects are rather well controlled: LIME relies heavily on simulations to generate data scenarios that are close to the data observed but exhibits variability. Effects from overplotting should be relatively mild, because output from simulations is shown, which is expected to be (relatively) continuous \hh{such that overplotting only occurs for points with (relatively) similar values.} \kgc{What do you mean by continuous here?}. \hh{XXX Does that extra explanation help?} \kgc{Yes. I see what you mean now.} In that respect, the diagnostics shown for the \data\ and in the bullet example are representative of what is expected. But in cases where overplotting does become problematic, the user could either simply reduce the size of the simulations or use some well-studied binning techniques in the visualizations, as discussed for example in \citet{carr:1987} or \citet{unwin:2006}.
